Understanding the Lomb-Scargle Periodogram

### 2016-12-27 PracticalLombScargle.tex
Add discussion of Classical->Lomb Scargle connection

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig13_nyquist_eyer99}
  \caption{A visualization of the \citep{Eyer99} definition of the Nyquist
    frequency. Data are non-uniformly sampled at times $t_i = n_i p$, for
    integer $n_i$ and $p=0.01$.
    This results in a Nyquist frequency of $f_{Ny}= (2p)^{-1} = 50$:
    the periodogram outside the range $0 \le f < f_{Ny}$ is a series of
    repetitions of the signal within that range.
    \figlabel{nyquist-eyer99}}
\end{figure}

@@

\citet{Scargle82} shows that the classical periodogram of
\eq{classical-periodogram} has such desireable statistical properties in the
case of uniform sampling, but not in the case of non-uniform sampling.
Such desireable properties can be recovered in the non-uniform case by
applying a slight modification to the standard periodogram:
\begin{equation}
  P(f)
  = \frac{1}{2}\left[
    \frac{\left[\sum_n g_n \cos(2\pi f [t_n-\tau])\right]^2}
         {\sum_n g_n \cos^2(2\pi f [t_n-\tau])}
  + \frac{\left[\sum_n g_n \sin(2\pi f [t_n-\tau])\right]^2}
         {\sum_n g_n \sin^2(2\pi f [t_n-\tau])}
  \right]
  \eqlabel{lomb-scargle-periodogram}
\end{equation}
where $\tau$ is given by
\begin{equation}
  \tan(4\pi f \tau) = \frac{\sum_n 4\pi f \tau}{\sum_n 4\pi f\tau}.
  \eqlabel{tau-def}
\end{equation}
This form has since become commonly referred to as the ``Lomb-Scargle Periodogram'', after \citet{Lomb76} and \citet{Scargle82} who proposed this alternate
periodogram and explored its statistical properties.

@@

\begin{itemize}
\item Bayesian view is (almost always) not useful
\item Multiterm makes the true period fit better, but also bumps the background noise.
\end{itemize}

>>>>

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig13_nyquist_eyer99}
  \caption{A visualization of the \citep{Eyer99} definition of the Nyquist
    frequency. Data are non-uniformly sampled at times $t_i = n_i p$, for
    integer $n_i$ and $p=0.01$.
    This results in a Nyquist frequency of $f_{Ny}= (2p)^{-1} = 50$:
    the periodogram outside the range $0 \le f < f_{Ny}$ is a series of
    perfect aliases of the signal within that range.
    \figlabel{nyquist-eyer99}}
\end{figure}

@@

\citet{Scargle82} showed that while the classical periodogram of
\eq{classical-periodogram} has these desireable statistical properties
in the case of uniform sampling, this does not hold for
\eq{classical-periodogram} when the sampling becomes non-uniform.
To address this, he considered a generalized form of the periodogram,
\begin{equation}
  P(f) = \frac{A^2}{2}\left(\sum_n g_n \cos(2\pi f t_n)\right)^2
       + \frac{B^2}{2} \left(\sum_n g_n \sin(2\pi f t_n)\right)^2,
\end{equation}
where $A$ and $B$ are arbitrary functions of the frequency $f$ and
observing times $\{t_i\}$ (but not the values $\{g_n\}$, and showed
that you can choose one particular form of $A$ and $B$ such that
\begin{itemize}
  \item The periodogram reduces to the classical form in the case of equally-spaced data,
  \item The periodogram has desireable statistical properties, in the sense discussed above,
  \item The periodogram is insensitive to global time-shifts in the data.
\end{itemize}
The values of $A$ and $B$ leading to these properties lead to the following
modification of the classical expression for the periodogram:
\begin{eqnarray}
  P(f) =
  \frac{1}{2} \Bigg\{ &
  \bigg(\sum_n g_n \cos(2\pi f [t_n-\tau])\bigg)^2 \bigg/
  \sum_n \cos^2(2\pi f [t_n-\tau]) &\nonumber\\
  & + ~ \bigg(\sum_n g_n \sin(2\pi f [t_n-\tau])\bigg)^2 \bigg/
  \sum_n \sin^2(2\pi f [t_n-\tau]) & \Bigg\}
  \eqlabel{lomb-scargle-periodogram}
\end{eqnarray}
Here $\tau$ is defined for each $f$ to ensure time-shift invariance of $P(f)$:
\begin{equation}
  \tau = \frac{1}{4\pi f}\tan^{-1}\Bigg(
  \frac{\sum_n \sin(4\pi f t_n)}{\sum_n \cos(4\pi f t_n)}\Bigg).
  \eqlabel{tau-def}
\end{equation}
A remarkable feature of this result is that it is {\it identical} to the result
obtained by fitting a simple sinusoid to the data at each frequency $f$, and
constructing a ``periodogram'' from the $\chi^2$ goodness-of-fit at each
frequency--an estimator which was considered in some depth by \citet{Lomb76}.
Partly due to this deep connection between Fourier analysis and least-squares
analysis, the modified periodogram in \eq{lomb-scargle-periodogram}
has since become commonly referred to as the ``Lomb-Scargle Periodogram''.

Often, the Lomb-Scargle periodogram only differs slightly from the
classical/Schuster periodogram; an example of this is seen in
\fig{ls-comparison}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig17_ls_comparison}
  \caption{A comparison of the Classical periodogram
    (\eq{classical-periodogram}) and the Lomb-Scargle periodogram
    (\eq{lomb-scargle-periodogram}) for 30 noisy points drawn from a sinusoid.
    Though the two differ quantitatively, the essential qualitative
    features---namely the position of significant peaks---generally remains
    the same.
    \figlabel{ls-comparison}}
\end{figure}

\subsection{Properties of the Lomb-Scargle Periodogram}

@@

\begin{itemize}
\item Uncertainties in Data
\item Bayesian view is (almost always) not useful
\item Multiterm makes the true period fit better, but also bumps the background noise.
\end{itemize}
